\documentclass[11pt,a4paper]{article}

% Alias pi name
\usepackage{xspace}
\newcommand{\rpi}{\textit{Raspberry Pi\textsuperscript{\textregistered}}}
\newcommand{\rpis}{\textit{Raspberry Pi\textsuperscript{\textregistered}s}}

% Condense list whitespace
\usepackage{enumitem}
\setitemize{noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}


\begin{document}

\section*{Operating system principles}

\section*{Team $\langle$sql injection$\rangle$}

Our project is a modernised phonograph, using the \rpi\xspace for subject tracking, digital recording and audio processing.

\subsection*{Learning Objective 1}


\begin{itemize}
  \item Prototyped complex code in Python, then translated to C/C++ for better performance.
  \item Used static analysers to find code improvements.
  \item Used ``Instruments" for debugging performance and finding memory leaks.
  \item Ran manual benchmarks to test performance and accuracy.
  \item Git and GitHub for version control.
\end{itemize}

\subsection*{Learning Objective 2}

\begin{itemize}
  \item Designed workload to be split across two \rpis.
  \item Iteratively improved CPU performance of face detection.
  \item Incorporated motion sensing to improve power consumption.
  \item Experimented with process scheduling.
\end{itemize}

\subsection*{Learning Objective 3}

\begin{itemize}
  \item Carefully designed a 19 page project specification.
  \item Documented message queue between two \rpis.
  \item Wrote a multithreaded solution.
  \item Experimented with process scheduling.
\end{itemize}

\subsection*{Learning Objective 4}

\begin{itemize}
  \item Modularised solution into weekly deliverables.
  \item Arranged deliverables so dependencies were not in sequential weeks.
  \item Wrote software to simulate hardware responses.
  \item Cross-skill development.
\end{itemize}

% PIR battery performance.

\end{document}